Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-23 NEUROC/sec4-experiments.tex

With the sampling techniques, GNNs were trained in a mini-batch manner.
%
Each epoch was decomposed into many batches.
%
In each batch, PyG only sent the sampled subgraph to the GPU to train or inference.
%
Since the sampled subgraphs could be much smaller than the original graph, the experiments in this section focused on finding out how the sample size affected the training time, the memory usage and the accuracy of GNNs.

@@

To find out performance bottlenecks of the sampling techniques, we decomposed the training time per batch into three phases: \emph{sampling} on the CPU side, \emph{transferring} sampled subgraphs from the CPU side to the GPU side, and \emph{training} with the sampled subgraphs on the GPU side.
%
\figurename~\ref{fig:exp_sampling_batch_train_time} shows the time breakdown of the four GNNs under different relative batch sizes.
%
For the neighbor sampler, the sampling technique reduced the training time per batch only when the batch size was very small.
%
When the batch became bigger, the sampling and the data transferring phases introduced noticeable overheads, making the training time exceed the full-batch training.
%
For the clustering sampler, the sampled subgraph was smaller than the neighbor sampler under the same relative batch size.
%
The reduction in the training time was more obvious than the neighbor sampler.
%
However, the overheads increased quickly as the relative batch size increased.
%
The training time under the 25\% relative batch size already exceeded the time of full-batch training.
%
The experimental results indicated that the current implementation of the sampling techniques in PyG was inefficient.
%
When the batch size was large, more than 50\% of the time had been spent on sampling and data transferring.
%
\emph{The sampling techniques were only efficient under small batch sizes.}

@@

% added by wangyunpan, 2020-12-23, ----start----
\paragraph{Effects on Accuracy}
%
We were also concerned about whether the accuracy would be significantly dropped after the use of sampling techniques
and the effect of batch size of accuracy.
%
For all datasets and sampling techniques, we had done experiments where the accuracy of GNNs varied with the batch size
and compared them with the experimental results of full-batch training.
%
In order to make the results more reliable, we had selected the optimal network structure for each GNN experiment under each dataset.
%
The optimal network structure is taken from the result corresponding to the highest accuracy in the hyperparameter experiment.
%
The highest accuracy under different learning rates was selected as the final accuracy result.
%
\figurename~\ref{fig:fig:exp_sampling_relative_batch_size_accuracy} shows the results (we only shows two datasets, amc and fli).
%

%
With the change of batch size, the accuracy relationship between GNNs is roughly the same.
%
For example, for neigbor sampler, the relationship between GNNs on \texttt{amc} was GCN < GGNN < GAT < GaAN and the relationship on \texttt{amc} was GCN > GGNN > GAT> GaAN.
%
But there was also a little exceptions (cluster sampler on \textt{amc}, relative batch size was 50\%).
%
To a certain extent, the increase of batch size could bring benefits to all GNNs.
%
However, for different datasets, the accuracy relationship between GNNs was not constant.
%
And this relationship was not related to model complexity.
%
In other words, it was not that the higher the algorithm complexity, the higher the corresponding accuracy.
%
For a specific dataset, in order to obtain high accuracy, the best model could only be selected by constantly trying various models.
%

%
After using the sampling techniques, the optimal accuracy of some GNNs on certain datasets exceeded the accuracy of the full-batch training (neigbbor sampler on \texttt{amc}).
%
This showed that the application of sampling technology was very meaningful, because it may brought certain benefits to accuracy.
%
Different sampling techniques bring different gains to GNNs.
%
\begin{figure}[H]
    \centering
    \subfloat[Neighbor sampler on \texttt{amc}]{\includegraphics[height=5cm]{figs/experiments/exp_graphsage_sampling_accuracy_on_amc.pdf}}
    \subfloat[Neighbor sampler on \texttt{fli}]{\includegraphics[height=5cm]{figs/experiments/exp_graphsage_sampling_accuracy_on_fli.pdf}} \\
    \subfloat[Cluster sampler on \texttt{amc}]{\includegraphics[height=5cm]{figs/experiments/exp_cluster_sampling_accuracy_on_amc.pdf}}
    \subfloat[Cluster sampler on \texttt{fli}]{\includegraphics[height=5cm]{figs/experiments/exp_cluster_sampling_accuracy_on_fli.pdf}}
    \caption{Test Accuracy under different batch sizes. FULL means that the full graph participates in the training.}
    \label{fig:exp_sampling_relative_batch_size_accuracy}
\end{figure}

>>>>

With the sampling techniques, GNNs were trained and inferenced in a mini-batch manner.
%
Each epoch was decomposed into many batches.
%
In each batch, PyG only sent the sampled subgraph to the GPU to train or inference.
%
The batch size determined the size of the sampled subgraph.
%
Thus, it affected the accuracy, training/inference time and memory usage of GNNs simultaneously.

\subsubsection{Size of Sampled Subgraphs}

@@

\subsubsection{Effects on Accuracy}

Since the size of sampled subgraph was much smaller than the original graph, it might affect the accuracy of the trained GNNs.
%
To find out how the batch size affected the accuracy, we trained GNNs under different batch sizes and compared their accuracy with the full-batch training in \figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy}.
%
The experiemental results on other datasets were similar.
%
We used the same hyper-parameters for the sample-based training and the full-batch training.
%
For each dataset and each GNN, we chose the hyper-parameters that achieved the highest accuracy for the full-batch training, according to the experimental results in Section~\ref{sec:effects_of_hyper-parameters_on_performance}.

\begin{figure}[H]
    \centering
    \subfloat[Neighbor sampler on \texttt{amc}]{\includegraphics[height=5cm]{figs/experiments/exp_graphsage_sampling_accuracy_on_amc.pdf}}
    \subfloat[Neighbor sampler on \texttt{fli}]{\includegraphics[height=5cm]{figs/experiments/exp_graphsage_sampling_accuracy_on_fli.pdf}} \\
    \subfloat[Cluster sampler on \texttt{amc}]{\includegraphics[height=5cm]{figs/experiments/exp_cluster_sampling_accuracy_on_amc.pdf}}
    \subfloat[Cluster sampler on \texttt{fli}]{\includegraphics[height=5cm]{figs/experiments/exp_cluster_sampling_accuracy_on_fli.pdf}}
    \caption{Test Accuracy under different batch sizes. FULL means that the full graph participates in the training.}
    \label{fig:exp_sampling_relative_batch_size_accuracy}
\end{figure}

%
With the change of batch size, the accuracy relationship between GNNs is roughly the same.
%
For example, for neigbor sampler, the relationship between GNNs on \texttt{amc} was GCN < GGNN < GAT < GaAN and the relationship on \texttt{amc} was GCN > GGNN > GAT> GaAN.
%
But there was also a little exceptions (cluster sampler on \texttt{amc}, relative batch size was 50\%).
%
To a certain extent, the increase of batch size could bring benefits to all GNNs.
%
However, for different datasets, the accuracy relationship between GNNs was not constant.
%
And this relationship was not related to model complexity.
%
In other words, it was not that the higher the algorithm complexity, the higher the corresponding accuracy.
%
For a specific dataset, in order to obtain high accuracy, the best model could only be selected by constantly trying various models.
%

%
After using the sampling techniques, the optimal accuracy of some GNNs on certain datasets exceeded the accuracy of the full-batch training (neigbbor sampler on \texttt{amc}).
%
This showed that the application of sampling technology was very meaningful, because it may brought certain benefits to accuracy.
%
Different sampling techniques bring different gains to GNNs.
%

\subsubsection{Performance Bottlenecks Analysis}

To find out performance bottlenecks of the sampling techniques, we decomposed the training/inference time per batch into three phases: \emph{sampling} on the CPU side, \emph{transferring} sampled subgraphs from the CPU side to the GPU side, and \emph{training/inference} with the sampled subgraphs on the GPU side.

\paragraph{Training Time}

\figurename~\ref{fig:exp_sampling_batch_train_time} shows the \emph{training} time breakdown of the four GNNs under different relative batch sizes.
%
For the neighbor sampler, the sampling technique reduced the training time per batch only when the batch size was very small.
%
When the batch became bigger, the sampling and the data transferring phases introduced noticeable overheads, making the training time exceed the full-batch training.
%
For the clustering sampler, the sampled subgraph was smaller than the neighbor sampler under the same relative batch size.
%
The reduction in the training time was more obvious than the neighbor sampler.
%
However, the overheads increased quickly as the relative batch size increased.
%
The training time under the 25\% relative batch size already exceeded the time of full-batch training.
%
The experimental results indicated that the current implementation of the sampling techniques in PyG was inefficient.
%
When the batch size was large, more than 50\% of the time had been spent on sampling and data transferring.
%
\emph{The sampling techniques were only efficient under small batch sizes.}

@@
